---
title: "BVS regression on simulated data"
author: "Yifan Zhou"
date: "12/1/2018"
output: 
  html_document: 
    toc: true
    toc_float: true
---

<style>
h2 { /* Header 2 */
  color: #2D4D56;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE, comment = NA, fig.align = 'center')
```

Load functions for data simulation and BVS regression:
```{r}
source("code/data_gen.R")
source("code/BVS.R")
source("code/BVS_util.R")
```

## Data initialization

Here, we randomly generate simluation data by following these steps:

* generate $X_{N\times P}: X_{ij} \stackrel{i.i.d}{\sim} N(0,1)$;   
* generate $A_{1\times P}$:  i) $A_j \stackrel{i.i.d}{\sim} NB(20,0.25)$,  ii) set the overall sparsity level to $0.2$,  iii) $A_j \leftarrow \text{log } (A_j+1)$;   
* fix the enrichment parameters $\alpha_1=0.5$, $\alpha_0=-2$;   
* sample $\gamma$ given $\alpha,A$;    
* sample $\beta$ given $\gamma,\sigma^2_{\beta}=1$ and $\epsilon$ given $\sigma^2_e=1$;   
* generate $y = X\beta + \epsilon$.

```{r}
data = data.gen(N=100,P=100,X.sd=1,sparse=0.2,alpha_true=c(0.5,-2),sigma2_e_true=1,sigma2_b_true=1)
X = data$X
y = data$y
A = data$A
beta_true = data$beta_true
```

Here are all the true signals in the simulated data-set:
```{r}
plot(beta_true, pch=16, xlab='covariate index', ylab='effect size')
points((1:length(beta_true))[beta_true!=0],beta_true[beta_true!=0],col='red',pch=16)
```

<br/>

## BVS regression

```{r}
res = BVS(y=y,X=X,A=A,niter=1000,burn=200)
```

Examine the posterior inclusion probabilities of all covariates, and compare them with the true signals:

```{r fig.width=10, fig.height=5}
par(mfrow=c(1,2))
BVS_plot.PIP(pip=res$PIP, beta_true=beta_true)
BVS_plot.PIP(pip=res$PIP, beta_true=beta_true, thres_b=0.2)
```

'True signals' (the red dots) are defined as covariates with non-zero effect sizes in the left plot, and  covariates with absolute effect sizes > 0.2 in the right plot.

Calculate the average posterior inclusion probability over the true 'non-signals' vs the true signals:
```{r}
res.stats = BVS_selection.stats(pip=res$PIP,beta_true=beta_true)
res.stats$pp
```

Calculate the precision and recall rates by comparing the true signals with the estimated ones:
```{r}
res.stats$overlap
cat(paste('Precision of BVS estimation:',round(res.stats$overlap[3]/res.stats$overlap[2],digits = 2),'\n'))
cat(paste('Recall of BVS estimation:',round(res.stats$overlap[3]/res.stats$overlap[1],digits = 2)))
```

<br/>

## Session information

```{r}
sessionInfo()
```